---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SELECTRdata

<!-- badges: start -->
[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
<!-- badges: end -->

SELECTRdata provides convenience functions for downloading raster and tabular
data used in the Spatially Explicit Load Enrichment Calculation Tool (SELECT).
By providing a SpatRaster object of the target watershed, functions are available
to download cropped:
- [National Land Cover Dataset](https://www.mrlc.gov/)
- [FEMA USA Structures](https://disasters.geoplatform.gov/USA_Structures/)
- [Census Blocks](https://tigerweb.geo.census.gov/tigerwebmain/TIGERweb_restmapservice.html)
- [TIGER County Boundaries](https://tigerweb.geo.census.gov/tigerwebmain/TIGERweb_restmapservice.html)
- [USDA Agricultural Census](https://www.nass.usda.gov/)

Sources to add:
- Point sources (via ECHO or echor)
- MS4 urbanized areas (via US Census)

## Installation

You can install the development version of SELECTRdata like so:

``` r
# FILL THIS IN! HOW CAN PEOPLE INSTALL YOUR DEV PACKAGE?
```

## Example

## MRLC National Land Cover Dataset

```{r example}
library(conflicted)
library(SELECTRdata)
library(terra)

## we need a template file, this is the thomsoncreek watershed in Texas
dem <- system.file("extdata", "thompsoncreek.tif", package = "SELECTR")
dem <- terra::rast(dem)

gpkg <- system.file("extdata", "thompsoncreek.gpkg", package = "SELECTR")
wbd <- terra::vect(gpkg, layer = "wbd")

dem <- terra::mask(dem, wbd,
                   filename = tempfile(fileext = ".tif"))


```


```{r}
## set the following GDAL options to connect to
## MRLC's AWS S3 bucket
set_gdal_config("AWS_NO_SIGN_REQUEST", "YES")

## download the NLCD file cropped to the extents of the watershed
nlcd <- SELECTRdata::download_nlcd(template = dem, 
                                   overwrite = TRUE,
                                   progress = 1)
plot(nlcd)
```


### FEMA US Buildings

```{r}
buildings <- download_buildings(template = dem)
plot(buildings)
```

### TIGER Counties

```{r}
cen_blocks <- download_census_blocks(dem, "2020")
plot(cen_blocks, "POP100")
```

```{r}
counties <- download_counties(dem)
plot(counties)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

