---
output: github_document
---

<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. -->
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SELECTRdata

<!-- badges: start -->
[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![R-CMD-check](https://github.com/TxWRI/SELECTRdata/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/TxWRI/SELECTRdata/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/TxWRI/SELECTRdata/branch/main/graph/badge.svg)](https://app.codecov.io/gh/TxWRI/SELECTRdata?branch=main)
<!-- badges: end -->

SELECTRdata provides convenience functions for downloading raster and tabular
data used in the Spatially Explicit Load Enrichment Calculation Tool (SELECT).
By providing a SpatRaster object of the target watershed, functions are available
to download cropped:
- [National Land Cover Dataset](https://www.mrlc.gov/)
- [FEMA USA Structures](https://disasters.geoplatform.gov/USA_Structures/)
- [Census Blocks](https://tigerweb.geo.census.gov/tigerwebmain/TIGERweb_restmapservice.html)
- [TIGER County Boundaries](https://tigerweb.geo.census.gov/tigerwebmain/TIGERweb_restmapservice.html)
- [USDA Agricultural Census](https://www.nass.usda.gov/)
- [EPA NPDES Permits](https://echo.epa.gov/)
- [U.S. Census Bureau Urbanized Areas](https://www.census.gov/programs-surveys/geography/guidance/geo-areas/urban-rural.html)

Sources to add:
- Point sources (via ECHO or echor)
- MS4 urbanized areas (via US Census)

## Installation

You can install the development version of SELECTRdata like so:

``` r
install.packages("SELECTRdata", repos = c("https://txwri.r-universe.dev", "https://cloud.r-project.org"))
```

## Example

## MRLC National Land Cover Dataset

```{r example}
library(SELECTRdata)
library(terra)

## we need a template file, this is the thomsoncreek watershed in Texas
dem <- system.file("extdata", "thompsoncreek.tif", package = "SELECTRdata")
dem <- terra::rast(dem)

gpkg <- system.file("extdata", "thompsoncreek.gpkg", package = "SELECTRdata")
wbd <- terra::vect(gpkg, layer = "wbd")

dem <- terra::mask(dem, wbd,
                   filename = tempfile(fileext = ".tif"))


```


```{r}
## set the following GDAL options to connect to
## MRLC's AWS S3 bucket
set_gdal_config("AWS_NO_SIGN_REQUEST", "YES")

## download the NLCD file cropped to the extents of the watershed
nlcd <- SELECTRdata::download_nlcd(template = dem, 
                                   overwrite = TRUE,
                                   progress = 1)
plot(nlcd)
plot(wbd, add = TRUE)
```


### FEMA US Buildings

```{r}
buildings <- download_buildings(template = dem)
plot(buildings)
plot(wbd, add = TRUE)
```

### U.S. Census Blocks

Includes housing unit and population data.

```{r}
cen_blocks <- download_census_blocks(dem, "2020")
plot(cen_blocks, "P0010001")
plot(wbd, col = "white", alpha = 0.5, add = TRUE)
```

### TIGER Counties

County boundaries cropped to coastlines.

```{r}
counties <- download_counties(dem)
plot(counties)
plot(wbd, add = TRUE)
```

## Urbanized Areas

U.S. Census designated urban areas from the 2020 U.S. Census.

```{r}
ua <- download_urban_areas(dem)
plot(wbd)
plot(ua, col = "red", alpha = 0.5, add = TRUE)
```


## NPDES Permits


```{r}
npdes <- download_NPDES_permits(dem)
plot(wbd)
plot(npdes, add = TRUE)

```


